---
title: "Post-ABS Testing"
author: "Kaija Gahm"
date: '2022-08-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(gifski)
library(ggraph)
library(here)
library(igraph)
source(here("modelFunction_rewiring.R"))
```

First, run the model.

```{r}
# Define parameters
N = 100
edge.prob <- 0.04
burn.in = 20
burn.out = 5
pm = 0.3
ps = 0.1
pa = 0.2
add00 = c(0.5, 10)
lose01 = 0.1
add10 = 0.05
lose11 = c(0.5, 0.5)
histMultiplier = 1.2
doRemoval = TRUE
```

```{r}
modelGraphs <- runModel(N = N, # Nodes in the network
                     edge.prob = edge.prob, 
                     burn.in = burn.in,
                     burn.out = burn.out,
                     pm = pm,
                     ps = ps, 
                     pa = pa,
                     add00 = add00, 
                     lose01 = lose01, 
                     add10 = add10,
                     lose11 = lose11, 
                     histMultiplier = histMultiplier,
                     doRemoval = doRemoval) %>%
  lapply(., function(x){
    igraph::graph_from_adjacency_matrix(x, mode = "undirected", add.colnames = "label")
  })
```

1. How does variation in degree emerge in this model?

Compute degree for each of the model networks.
```{r}
degreesDF <- lapply(modelGraphs, function(x){
  set_vertex_attr(x, name = "degree", value = degree(x)) %>%
    tidygraph::as_tbl_graph() %>%
    tidygraph::activate(nodes) %>%
    data.frame()
  }) %>%
  data.table::rbindlist(idcol = "slice") %>%
  as.data.frame() %>%
  mutate(beforeAfter = ifelse(slice <= burn.in, "before", "after"))
```
Okay, so we get some degree distributions, but do we get emergent label variation in degree that's somewhat consistent throughout time?

```{r}
# Plot everyone's degree over time
degreesDF %>%
  ggplot(aes(x = slice, y = degree, col = label, group = label))+
  geom_line(alpha = 0.7, size = 0.5)+
  theme_minimal()+
  geom_vline(xintercept = burn.in+1, col = "red", size = 2)+
  theme(legend.position = "none")

# Let's just look at a few labels, because this plot is too hard to read
indivs <- sample(degreesDF$label, 5)
degreesDF %>%
  filter(label %in% indivs) %>%
  ggplot(aes(x = slice, y = degree, col = factor(label), group = label))+
  geom_line()+
  theme_minimal()+
  geom_vline(xintercept = burn.in+1, col = "red", size = 2)

# This is still a little hard to see... let's look at just the "before" rows.
degreesDF %>%
  filter(label %in% indivs, beforeAfter == "before") %>%
  ggplot(aes(x = slice, y = degree, col = factor(label), group = label))+
  geom_line()+
  theme_minimal()

# Okay, I'm seeing some differentiation, but no clear patterns. Because of how the model is set up, the rank order tends to be maintained for a few consecutive time steps, but not over the whole span of the baseline model. This makes sense, since labels don't have any a priori reason to have higher or lower degrees than other labels.
```

Now, going to compute three network-level measures before and after the removal of an label.

### 1. Network density deltas

How does edge density behave over time?
```{r}
densities <- lapply(modelGraphs, function(x){
  edge_density(x)
 }) %>% 
  unlist() %>% 
  as.data.frame() %>%
  setNames(., "density") %>%
  mutate(slice = 1:length(modelGraphs))

densities %>%
  ggplot(aes(x = slice, y = density))+
  geom_line()+
  geom_smooth()+
  theme_minimal()+
  geom_vline(aes(xintercept = burn.in+1), size = 1, col = "red")
```

First delta: immediately before removal to after removal/pre-rewiring
```{r}
delta1_density <- igraph::edge_density(modelGraphs[[burn.in+1]])-igraph::edge_density(modelGraphs[[burn.in]])
```

Second delta: after removal/pre-rewiring to post-rewiring
```{r}
delta2_density <- igraph::edge_density(modelGraphs[[burn.in+2]])-igraph::edge_density(modelGraphs[[burn.in+1]])
```

### 2. Network connectivity (average path length) deltas

How does connectivity behave over time?
```{r}
connectivities <- lapply(modelGraphs, function(x){
  mean_distance(x)
 }) %>% 
  unlist() %>% 
  as.data.frame() %>%
  setNames(., "connectivity") %>%
  mutate(slice = 1:length(modelGraphs))

connectivities %>%
  ggplot(aes(x = slice, y = connectivity))+
  geom_line()+
  geom_smooth()+
  theme_minimal()+
  geom_vline(aes(xintercept = burn.in+1), size = 1, col = "red")
```

First delta: immediately before removal to after removal/pre-rewiring
```{r}
delta1_meanDistance <- igraph::mean_distance(modelGraphs[[burn.in+1]])-igraph::mean_distance(modelGraphs[[burn.in]])
```

Second delta: after removal/pre-rewiring to post-rewiring
```{r}
delta2_meanDistance <- igraph::mean_distance(modelGraphs[[burn.in+2]])-igraph::mean_distance(modelGraphs[[burn.in+1]])
```

### 3. Network modularity deltas

How does modularity behave over time?
```{r}
modularities <- lapply(modelGraphs, function(x){
  cluster_fast_greedy(x) %>%
    modularity()
 }) %>% 
  unlist() %>% 
  as.data.frame() %>%
  setNames(., "modularity") %>%
  mutate(slice = 1:length(modelGraphs))

modularities %>%
  ggplot(aes(x = slice, y = modularity))+
  geom_line()+
  geom_smooth()+
  theme_minimal()+
  geom_vline(aes(xintercept = burn.in+1), size = 1, col = "red")
```

First delta: immediately before removal to after removal/pre-rewiring
```{r}
delta1_modularity <- modularity(cluster_fast_greedy(modelGraphs[[burn.in+1]]))-modularity(cluster_fast_greedy(modelGraphs[[burn.in]]))
```

Second delta: after removal/pre-rewiring to post-rewiring
```{r}
delta2_modularity <- modularity(cluster_fast_greedy(modelGraphs[[burn.in+2]]))-modularity(cluster_fast_greedy(modelGraphs[[burn.in+1]]))
```

## What should our measure be?
I am not at all convinced that these momentary deltas are what I should be using. Seems like we need to at least consider two time steps out, since the model relies on this. Looking at the graphs above shows me how different the longer-term trends are from the short-term fluctuations.

Let's run the model 100 times and see if trends emerge in how to measure.
```{r}
nRuns <- 100
modelRuns <- vector(mode = "list", length = nRuns)
for(i in 1:nRuns){
  modelRuns[[i]] <- runModel(N = N, # Nodes in the network
                     edge.prob = edge.prob, 
                     burn.in = burn.in,
                     burn.out = burn.out,
                     pm = pm,
                     ps = ps, 
                     pa = pa,
                     add00 = add00, 
                     lose01 = lose01, 
                     add10 = add10,
                     lose11 = lose11, 
                     histMultiplier = histMultiplier,
                     doRemoval = doRemoval) %>%
  lapply(., function(x){
    igraph::graph_from_adjacency_matrix(x, mode = "undirected")
  })
}
```

# Compute the three network-level measures for each of the model runs.
```{r}
densList <- lapply(modelRuns, function(x){
  lapply(x, function(y){
    edge_density(y)
  }) %>%
    unlist() %>%
    as.data.frame() %>%
    setNames(., "density") %>%
    mutate(slice = 1:nrow(.))
})
densDF <- data.table::rbindlist(densList, idcol = "run")

connList <- lapply(modelRuns, function(x){
  lapply(x, function(y){
    mean_distance(y)
  }) %>%
    unlist() %>%
    as.data.frame() %>%
    setNames(., "connectivity") %>%
    mutate(slice = 1:nrow(.))
})
connDF <- data.table::rbindlist(connList, idcol = "run")

moduList <- lapply(modelRuns, function(x){
  lapply(x, function(y){
    cluster_fast_greedy(y) %>%
    modularity()
  }) %>%
    unlist() %>%
    as.data.frame() %>%
    setNames(., "modularity") %>%
    mutate(slice = 1:nrow(.))
})
moduDF <- data.table::rbindlist(moduList, idcol = "run")
```

Make some plots and try to detect any changepoint trends
```{r}
densDF %>%
  ggplot(aes(x = slice, y = density))+
  geom_line(aes(group = run, col = run), size = 0.1)+
  geom_smooth()+
  theme_minimal()+
  geom_vline(aes(xintercept = burn.in+1), size = 0.5, col = "red")+
  scale_color_viridis_c()

connDF %>%
  ggplot(aes(x = slice, y = connectivity))+
  geom_line(aes(group = run, col = run), size = 0.1)+
  geom_smooth()+
  theme_minimal()+
  geom_vline(aes(xintercept = burn.in+1), size = 0.5, col = "red")+
  scale_color_viridis_c()

moduDF %>%
  ggplot(aes(x = slice, y = modularity))+
  geom_line(aes(group = run, col = run), size = 0.1)+
  geom_smooth()+
  theme_minimal()+
  geom_vline(aes(xintercept = burn.in+1), size = 0.5, col = "red")+
  scale_color_viridis_c()
```
